---
title: "Introduction to Data Science"
author: "Austin Adams  - `aadams0230@floridapoly.edu`"
subtitle: Homework 2
output:
  html_document:
    df_print: paged
---



# Problem 1 

(a) (10 points) Define a function `imperial_to_metric` that takes in two arguments: a number of feet and a number of inches. The function should return the equivalent length in meters
```{r}
imperial_to_metric <- function(var1,var2) #feet and inches
  {
 metric <- (var1+(var2/12))*0.3048 # takes total feet length and multiplies it by the conversion to get the metric number. 
 return(metric)
}
```



(b) (5 points) Create a variable `height_in_meters` by passing your height in imperial to the `imperial_to_metric` function

```{r}
height_in_meters <- imperial_to_metric(6,6) #this indicates someone who is 6 foot 6 inch
print(height_in_meters)
```


# Problem 2

`R` comes with many functions that you can use to do sophisticated tasks like random sampling. For example, you can round a number with the `round()` function, or calculate its factorial with the `factorial()` function. Using a function is pretty simple. Just write the name of the function and then the data you want the function to operate on in parentheses.

(a) (5 points) Find `round(3.1415)`, `factorial(12)`, `mean(1:17)`.
```{r}
print(round(3.1315))
print(factorial(12))
print(mean(1:17))
```


(b) (5 points) You can create an object named `die` that contains the numbers one through six:

```{r}
# Create a vector with numbers from 1 to 6
die <- 1:6
```

You can simulate a roll of the die with `R`’s `sample()` function. `sample()` takes two arguments: a vector named `x` and a number named `size`. 

Try `sample(x = die, size = 1)`. 
```{r}
sample(x=die,size = 1)
```

Is there any difference if you type `sample(die, size = 1)` ?
```{r}
sample(die, size=1)
```
_There is not any differences in the output with this change._


# Problem 3

Use `R` as a calculator to compute the following values:

(a) (5 points) $27 (38 -17)$
```{r}
(27*(38 -17))
```


(b) (5 points) $ln(14^7)$
```{r}
log(14^7)
```

(c) (5 points) $\sqrt\frac{436}{12}$

```{r}
sqrt(436/12)
```


# Problem 4

Create the following vectors in `R`:

$a = (5, 10, 15, 20, ..., 160)$
```{r}
a <- seq(5,160,5)
print(a)
```


$b = (87, 86, 85, ..., 56)$
```{r}
b <- 87:56
print(b)
```


(a) (5 points) Multiply these vectors and call the result `d`. 
```{r}
d <- a*b
print(d)
```


(b) (15 points) Select subsets of `d` to identify the following.

- What are the 19th, 20th, and 21st elements of `d`?
```{r}
d[19:21]
```


- What are all of the elements of `d` which are less than 2000?
```{r}
less_than <- d[d<2000]
print(less_than)
```

- How many elements of `d` are greater than 6000?
```{r}
greater_than <- d[d > 6000]
print(greater_than)
```



# Problem 5

Start with this vector of `nouns`.  

```{r}
nouns <- c("apple", "flower", "insect", "lettuce", "knife", "dog", "cloud", "person", "cabinet", "flower" )
```

(a) (5 points) Use the `length()` function to display the number of elements in `nouns`.
```{r}
length(nouns)
```


(b) (5 points) Use indexing to create a new vector consisting of the first 4 elements of `nouns`.
```{r}
nouns4 <- nouns[1:4]
print(nouns4)
```


(c) (5 points) Use indexing to create a new vector consisting of only the last 8 elements of `nouns.
```{r}
nouns[(length(nouns)-7):length(nouns)]
```


(d) (5 points) Use indexing to create a new vector of the 1st, 3rd through 6th, and 10th elements of `nouns` (the length of the resulting vector should be 6)
```{r}
indices <- c(1,3:6,10)
jumbled <- nouns[indices]
print(jumbled)
```



(e) (5 points) Create a new vector with the elements of `nouns` in reverse order.
```{r}
reversenouns <- rev(nouns)
reversenouns
```

# Problem 6

Consider the function `rescale01` defined as 

```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

(a) (5 points) Evaluate `rescale01(c(0, 5, 10))`
```{r}
rescale01(c(0,5,10))

```


(b) (5 points) Describe what the `rescale01` function does.

_It is re-scaling the vector entered to be a range between 0 and 1. The max value entered goes in as 1 and the min value entered goes in as 0. The middle number is within that range._

(c) (5 points) What would happen if `x` contained a single missing value, and `na.rm` was `FALSE`? Show an example.
```{r}
rescale01(c(NA, 5, 10))
```
_It takes the two values and makes them the respective max and min values. Resulting in the largest number being the max value and the lowest number the min._


